Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2018-02-07 LaTeX/main.tex
striking image added

If haploid selection is strong but selective differences between male and female diploids are weak, we find that trans-GSD transitions are favoured more strongly than cis-GSD transitions, with transitions to ESD intermediate (e.g., with $|D^\male - D^\female| << |\alpha_\Delta^\male - \alpha_\Delta^\female + t^\male - t^\female|$ we have $\lambda_{W'}^{(XY)} > \lambda_{Y'}^{(XY)}$; Equations \ref{eq:lambda_neoW} and \ref{eq:freq_diffs}). 
Among the relatively few dioecious clades in which multiple species have well characterized sex chromosomes~\cite{Ming:2011iy}, trans-GSD transitions have been inferred in \textit{Silene} subsection \textit{Otites}~\cite{Slancarova:2013dq} and in \textit{Salicaceae}~\cite{Pucholt2015,Pucholt2017}.
%Potentially, successive heterogametic transitions between master regulators of sex-determination could be inferred from careful examination of the molecular pathways by which sex is determined. 
%Our predictions could also be examined using a suitable proxy for haploid selection, for example, \cite{Lenormand:2005vb} use the outcrossing rate in plants as a proxy for the strength of pollen competition. 
Assuming that transitions from dioecy to hermaphroditism (equal parental investment in male and female gametes) are favoured in a similar manner to the ESD examined here (equal probability of zygotes developing as males or females), our results suggest that competition during the haploid stage could also drive transitions between dioecy and hermaphroditism, which are frequent in plants~\cite{Kafer2017, Goldberg2017}.

>>>>

If haploid selection is strong but selective differences between male and female diploids are weak, we find that trans-GSD transitions are favoured more strongly than cis-GSD transitions, with transitions to ESD intermediate (e.g., with $|D^\male - D^\female| << |\alpha_\Delta^\male - \alpha_\Delta^\female + t^\male - t^\female|$ we have $\lambda_{W'}^{(XY)} > \lambda_{Y'}^{(XY)}$; Eqs \ref{eq:lambda_neoW} and \ref{eq:freq_diffs}). 
Among the relatively few dioecious clades in which multiple species have well characterized sex chromosomes~\cite{Ming:2011iy}, trans-GSD transitions have been inferred in \textit{Silene} subsection \textit{Otites}~\cite{Slancarova:2013dq} and in \textit{Salicaceae}~\cite{Pucholt2015,Pucholt2017}.
%Potentially, successive heterogametic transitions between master regulators of sex-determination could be inferred from careful examination of the molecular pathways by which sex is determined. 
%Our predictions could also be examined using a suitable proxy for haploid selection, for example, \cite{Lenormand:2005vb} use the outcrossing rate in plants as a proxy for the strength of pollen competition. 
Assuming that transitions from dioecy to hermaphroditism (equal parental investment in male and female gametes) are favoured in a similar manner to the ESD examined here (equal probability of zygotes developing as males or females), our results suggest that competition during the haploid stage could also drive transitions between dioecy and hermaphroditism, which are frequent in plants~\cite{Kafer2017, Goldberg2017}.
